Evaluating a simplified forecast model in comparison to the European forecasting hub ensemble

### 2022-10-07 paper/paper.Rmd
add table

@@

Based on these insights, here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@Bosse2022-hm] to produce an ensemble-like performance without sacrificing interpretability.
The first simplification we make is to model only weekly data, rather than using daily data and then aggregating. This mitigates the impact of daily reporting artefacts. It also serves to increase the auto-correlation of the forecasting model as there is an increased lag before changes in daily observations gain significant weight in the model.
This leads to the observed ensemble behaviour of being relatively auto-correlated and resistant to short-term changes in trend.

@@

The final simplification is to model the growth rate as a differenced auto-regressive process with order 1 rather than using a gaussian process-based method as we have done in other work [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
This represents a parsimonious approach in that we encode our expectation that the growth rate should vary over time and allow this to influence the forecast but we include only a single lag term, reducing the computational overhead of the model. 
To model potential unobserved interventions and more general changes in transmission, we include an additional growth rate modifier restricted to be between 0 and 1 that differs depending on if the growth rate is positive or negative (due to potential differing responses when cases are growing or increasing) and that acts in a multiplicative fashion (meaning that larger absolute growth rates are reduced to zero growth more rapidly).
This reflects a simplified interpretation of how the ensemble appears to react to potential future changes by assuming a gradual return to stable incidence.

The only observation for which we do not make an adaptation is the apparent sharpness of the ensemble compared to our prior forecasting models choosing instead to make use of a negative binomial observation model allowing the inclusion of overdispersion.
This choice is motivated by our belief that the underlying transmission process is an exponential discrete one and therefore a count error model with a log link function, where variance is linked to the mean, is a sensible choice. 
We suggest that part of the reason the hub ensembles exhibit this behaviour is due to the penalisation of overprediction compared to underprediction caused by the use of a generalised form of absolute error for the majority of forecast evaluations [@bracherEvaluatingEpidemicForecasts2021].

>>>>

```{r}
observations <- data.table(
  No. = 1:5,
  Observation = c(
    "Robust to daily reporting artefacts",
    "Some ability to forecast future trend changes",
    "Less reactive to apparent observed changes in trend",
    "Sharper forecasts",
    "A tendency towards under prediction"
  )
)

assumptions <- data.table(
  Assumption = c(
    "Reported cases can be modelled using weekly data and a generative process discretised by week", # nolint
    "Reported cases can be modelled as if they represented infections",
    "The growth rate of infections can be represented as an auto-regressive process with an order of 1 week", # nolint
    "Unobserved interventions and more general changes in transmission towards a stable state can be represented using a multiplicative decay parameter" # nolint
  ),
  Observation = c("1, and 2", "-", "3 and 4", "2, and 5")
)
knitr::kables(
  list(knitr::kable(observations), knitr::kable(assumptions)),
  caption = "(\\#tab:observations) Observations on the relative performance of the Forecast Hub ensemble compared to our forecast submissions and assumptions/simplifications based on these observations." # nolint
)
```

@@

Based on these insights (summarised in Table \@ref(tab:observations)), here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@Bosse2022-hm] to produce an ensemble-like performance without sacrificing interpretability.
The first simplification we make is to model only weekly data, rather than using daily data and then aggregating. This mitigates the impact of daily reporting artefacts. It also serves to increase the auto-correlation of the forecasting model as there is an increased lag before changes in daily observations gain significant weight in the model.
This leads to the observed ensemble behaviour of being relatively auto-correlated and resistant to short-term changes in trend.

@@

The final simplification is to model the growth rate as a differenced auto-regressive process with an order 1 rather than using a gaussian process-based method as we have done in other work [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
This represents a parsimonious approach in that we encode our expectation that the growth rate should vary over time and allow this to influence the forecast but we include only a single lag term, reducing the computational overhead of the model. 
To model potential unobserved interventions and more general changes in transmission, we include an additional growth rate modifier restricted to be between 0 and 1 that differs depending on if the growth rate is positive or negative (due to potential differing responses when cases are growing or increasing) and that acts in a multiplicative fashion (meaning that larger absolute growth rates are reduced to zero growth more rapidly).
This reflects a simplified interpretation of how the ensemble appears to react to potential future changes by assuming a gradual return to stable incidence.

The only observation for which we do not make an adaptation is the apparent sharpness of the ensemble compared to our prior forecasting models choosing instead to make use of a negative binomial observation model allowing the inclusion of overdispersion.
This choice is motivated by our belief that the underlying transmission process is an exponential discrete one and therefore a count error model with a log link function, where variance is linked to the mean, is a sensible choice. 
We suggest that part of the reason the hub ensembles exhibit this behaviour is due to the penalisation of overprediction compared to underprediction caused by the use of a generalised form of absolute error for the majority of forecast evaluations [@bracherEvaluatingEpidemicForecasts2021]. Our derived set of assumptions and simplifications is summarised in Table \@ref(tab:observations).
